---
title: "Useful tricks in R"
output:  github_document
---
```{r}
library(dplyr)
library(tidyr)
```

```{r}
df<-read.csv('1_Intro/AnonymityPoll.csv')
```

### El classico : compute mean on columns and getting an **NA**

```{r}
mean(df$Internet.Use)
#ADD na.rm=TRUE
mean(df$Info.On.Internet,na.rm = TRUE)
```


### Get the **number of na per columns**
```{r}
df%>% summarise_each(funs(sum(is.na(.)))) %>% gather()
```

### Convert the index into column (useful with correlation matrix e.g)
```{r}
tibble::rownames_to_column(data.frame(df),'indexo')
```

### Get the number of NA per row :
```{r}
head(rowSums(is.na(df)))
```


### Get the **ratio of NA** per columns :
```{r}
colMeans(is.na(df))
```

### Convert a model summary into a data.frame :
```{r}
require(broom)
model1<-lm(Age~.,data=df %>% select_if(is.numeric))
df_model_summary<-tidy(model1)
head(df)
```


### Get the max/min of each columns :
```{r}
max_min_cols<-df %>%  select() summarise_each(funs(max,min))
head(max_min_cols %>% gather(),5)
```


### Select columns based on vector of string :
```{r}
select_(df2, .dots = vector_of_columns_name)

```

### Extract first part of string in vector :
```{r}
vapply(strsplit(high_skewed_features,"_"), `[`, 1, FUN.VALUE=character(1))
```

### Plot a lot of data on the same grid :
```{r}
library(gridExtra)

plot_data_column = function (data, column){
      p<-ggplot(data = data, aes_string(x = column)) +
        geom_histogram(fill = "lightgreen",binwidth = .1) +
        xlab(column)
}

myplots <- lapply(high_skewed_features, plot_data_column, data = df_skewed_features_log)


n <- length(myplots)
nCol <- floor(sqrt(n))
do.call("grid.arrange", c(myplots, ncol=nCol))

```

### Plot a correlogramm
```{r}
library(ggplot2)
library(ggcorrplot)
corr <- round(cor(ames_train_selected_features),1)

# Plot
ggcorrplot(corr, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 2, 
           method="square", 
           colors = c("cyan4", "white", "red3"), 
           title="Correlogram of Selected Features", 
           ggtheme=theme_bw)
```

### Combine multiple columns pair-wise :
```{r}

colos_Exp<-grep('X..Voix.Exp',colnames(df_votes.2017),value = T)
colos_name<-grep('Nom',colnames(df_votes.2017),value = T)
colos_Voix<-grep('^Voix',colnames(df_votes.2017),value = T)


filtered_df_pres<-df_votes.2017 %>% select_(.dots = c(colos_Exp,colos_name,colos_Voix,"Code.de.la.commune", "Libellé.de.la.commune","Code.du.département","Libellé.du.département","Votants"))

gather_name<-filtered_df_pres %>% gather(key,'Nom',starts_with('Nom')) 
gather_scores<- filtered_df_pres%>% gather(key,'Score',starts_with('X..Voix.Exp'))
gather_voix<- filtered_df_pres%>% gather(key,'Voix',starts_with('Voix'))


gather_name_scores<- data.frame(select(gather_scores, Score), select(gather_name,Nom,Code.de.la.commune, Libellé.de.la.commune,Code.du.département,Libellé.du.département,Votants),select(gather_voix,Voix))
```
 
  
## Summary of dataframe per class :
```{r}
print (table(sapply(df, class)))
```




