---
title: "Assignment 1 _An Analytical Detective"
output: html_notebook:keep_md: true
---


Problem 1.1 - Loading the Data
1 point possible (graded)
Read the dataset mvtWeek1.csv into R, using the read.csv function, and call the data frame "mvt". Remember to navigate to the directory on your computer containing the file mvtWeek1.csv first. It may take a few minutes to read in the data, since it is pretty large. Then, use the str and summary functions to answer the following questions.

How many rows of data (observations) are in this dataset?



```{r}
require(dplyr)
require(tidyr)

df <- read.csv('mvtWeek1.csv')
dim(df)
```
Problem 1.2 - Loading the Data
1 point possible (graded)
How many variables are in this dataset?
```{r}
str(df)
```

Problem 1.3 - Loading the Data
1 point possible (graded)
Using the "max" function, what is the maximum value of the variable "ID"?

```{r}
max(df$ID)
```
Problem 1.4 - Loading the Data
1 point possible (graded)
What is the minimum value of the variable "Beat"?
```{r}
min(df$Beat
    )
```
Problem 1.5 - Loading the Data
1 point possible (graded)
How many observations have value TRUE in the Arrest variable (this is the number of crimes for which an arrest was made)?
```{r}
table(df$Arrest)
```
Problem 1.6 - Loading the Data
1 point possible (graded)
How many observations have a LocationDescription value of ALLEY?
```{r}
sum(df$LocationDescription=='ALLEY')
```
Problem 2.1 - Understanding Dates in R
1 point possible (graded)
In many datasets, like this one, you have a date field. Unfortunately, R does not automatically recognize entries that look like dates. We need to use a function in R to extract the date and time. Take a look at the first entry of Date (remember to use square brackets when looking at a certain entry of a variable).

In what format are the entries in the variable Date?
```{r}
df$Date[nrow(df)/2]
```
Problem 2.2 - Understanding Dates in R
1 point possible (graded)
Now, let's convert these characters into a Date object in R. In your R console, type

DateConvert = as.Date(strptime(mvt$Date, "%m/%d/%y %H:%M"))

This converts the variable "Date" into a Date object in R. Take a look at the variable DateConvert using the summary function.

What is the month and year of the median date in our dataset? Enter your answer as "Month Year", without the quotes. (Ex: if the answer was 2008-03-28, you would give the answer "March 2008", without the quotes.)
```{r}
DateYear = as.Date(strptime(df$Date,"%m/%d/%y %H:%M"))
summary(DateYear)
median(DateYear)
```
Problem 2.3 - Understanding Dates in R
1 point possible (graded)
Now, let's extract the month and the day of the week, and add these variables to our data frame mvt. We can do this with two simple functions. Type the following commands in R:

mvt$Month = months(DateConvert)

mvt$Weekday = weekdays(DateConvert)

This creates two new variables in our data frame, Month and Weekday, and sets them equal to the month and weekday values that we can extract from the Date object. Lastly, replace the old Date variable with DateConvert by typing:

mvt$Date = DateConvert

Using the table command, answer the following questions.

In which month did the fewest motor vehicle thefts occur?
```{r}
df$Month = months(DateYear)
df$Weekday = weekdays(DateYear)
sort(table(df$Month),decreasing = F)
```
Problem 2.4 - Understanding Dates in R
1 point possible (graded)
On which weekday did the most motor vehicle thefts occur?
```{r}
sort(table(df$Weekday),decreasing = T)
```
Problem 2.5 - Understanding Dates in R
1 point possible (graded)
Each observation in the dataset represents a motor vehicle theft, and the Arrest variable indicates whether an arrest was later made for this theft. Which month has the largest number of motor vehicle thefts for which an arrest was made?
```{r}
table1<-table(df$Month,df$Arrest)
table1
which.max(table1[,2])
```

```{r}
df_p<-tbl_df(df)
groupy <- df_p %>% group_by(Month,Arrest) %>% tally() %>% spread(Arrest, n, fill=0)

groupy<-groupy%>%rowwise()%>%mutate(ratio = `TRUE`/sum(`FALSE`,`TRUE`)*100)
groupy %>% arrange(desc(`TRUE`)) %>% filter(row_number()==1 | row_number()==n())
```

