---
title: "Predicting Loan repayment"
output:
  html_document:
    keep_md: true
---

```{r}
library(dplyr)
library(tidyr)
library(caret)
library(ROCR)
library(mice)
library(broom)
df<- read.csv('loans.csv')

summary(df)

```
Problem 1.1 - Preparing the Dataset
1 point possible (graded)
Load the dataset loans.csv into a data frame called loans, and explore it using the str() and summary() functions.

What proportion of the loans in the dataset were not paid in full? Please input a number between 0 and 1.
```{r}
df %>% group_by(not.fully.paid) %>% summarise(total = n( )) %>% mutate(ratio= total*100/sum(total))
```
Problem 1.2 - Preparing the Dataset
1 point possible (graded)
Which of the following variables has at least one missing observation? Select all that apply.
```{r}
df%>% summarise_each(funs(sum(is.na(.)))) %>% gather() %>% filter(value>0)

```
Problem 1.3 - Preparing the Dataset
1 point possible (graded)
Which of the following is the best reason to fill in the missing values for these variables instead of removing observations with missing data? (Hint: you can use the subset() function to build a data frame with the observations missing at least one value. To test if a variable, for example pub.rec, is missing a value, use is.na(pub.rec).)
```{r}
not_fully_paid_ratio_only_no_NA <-df %>% na.omit() %>% summarise(mean(not.fully.paid))
not_fully_paid_ratio_only_with_NA <- df %>% mutate(has_NA = rowSums(is.na(.))) %>% filter(has_NA>0) %>% summarise(mean(not.fully.paid))

print(not_fully_paid_ratio_only_no_NA)
print(not_fully_paid_ratio_only_with_NA)

df %>% mutate(has_NA = rowSums(is.na(.))) %>% filter(has_NA>0) %>% nrow()
```


Problem 1.4 - Preparing the Dataset
1 point possible (graded)
For the rest of this problem, we'll be using a revised version of the dataset that has the missing values filled in with multiple imputation (which was discussed in the Recitation of this Unit). To ensure everybody has the same data frame going forward, you can either run the commands below in your R console (if you haven't already, run the command install.packages("mice") first), or you can download and load into R the dataset we created after running the imputation: loans_imputed.csv.

IMPORTANT NOTE: On certain operating systems, the imputation results are not the same even if you set the random seed. If you decide to do the imputation yourself, please still read the provided imputed dataset (loans_imputed.csv) into R and compare your results, using the summary function. If the results are different, please make sure to use the data in loans_imputed.csv for the rest of the problem.
```{r}
df2<- complete(mice(select(df,-not.fully.paid)))
summary(df2)
df2$not.fully.paid<-df$not.fully.paid
str(df2)
```
```{r}
set.seed(144)
library(caTools)

split = sample.split(df2$not.fully.paid, SplitRatio = 0.7)

df_train = subset(df, split == TRUE)

df_test= subset(df, split == FALSE)
```

```{r}

#df_train$purpose = relevel(df_train$purpose,'debt_consolidation')
model1 <- glm(not.fully.paid~.,data=df_train,family = binomial)
summary(model1)
```
```{r}
sumo<-tidy(model1)
rownames(sumo)<- sumo$term
str(sumo)
sumo %>% select(term,p.value) %>% filter(p.value<.05) %>% arrange(p.value)
``` 
Problem 2.2 - Prediction Models
4.0 points possible (graded)
Consider two loan applications, which are identical other than the fact that the borrower in Application A has FICO credit score 700 while the borrower in Application B has FICO credit score 710.

Let Logit(A) be the log odds of loan A not being paid back in full, according to our logistic regression model, and define Logit(B) similarly for loan B. What is the value of Logit(A) - Logit(B)?


  unanswered  
Now, let O(A) be the odds of loan A not being paid back in full, according to our logistic regression model, and define O(B) similarly for loan B. What is the value of O(A)/O(B)? (HINT: Use the mathematical rule that exp(A + B + C) = exp(A)*exp(B)*exp(C). Also, remember that exp() is the exponential function in R.)

```{r}
A<-700
B<-710
sumo['fico','estimate']*(A-B)

```
$$\begin{aligned}
Logit(A) =& log(Odds(A))\\\\
Logit(B) =& log(Odds(B)) \\\\
\frac{Odds(A)}{Odds(B)} =& \frac{e^{log(Odds(A))}}{e^{log(Odds(B))}}\\\\
\frac{Odds(A)}{Odds(B)} =& e^{log(Odds(A) - log(Odds(B)}\\\\
\frac{Odds(A)}{Odds(B)} =& e^{Logit(A)- Logit(B)}
\end{aligned}$$
Problem 2.3 - Prediction Models
4.0 points possible (graded)
Predict the probability of the test set loans not being paid back in full (remember type="response" for the predict function). Store these predicted probabilities in a variable named predicted.risk and add it to your test set (we will use this variable in later parts of the problem). Compute the confusion matrix using a threshold of 0.5.

What is the accuracy of the logistic regression model? Input the accuracy as a number between 0 and 1.


  unanswered  
What is the accuracy of the baseline model? Input the accuracy as a number between 0 and 1.
